---
title: "Lineage calling with SLIMM"
output:
    rmarkdown::html_vignette:
        toc: true
        toc_depth: 3
vignette: >
  %\VignetteIndexEntry{pkgdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Lineage calling

One thing you might want to do with metagenomics data is to identify the species
present in your sample. Here we will use [SLIMM](https://github.com/seqan/slimm)
which is a binning based classifier similar to Kraken.

The reason we often use SLIMM is that it generates cleaned species-level coverage
maps as a side effect which we use extensively.

```{r}
library(mbtools)
```

## Getting the SLIMM DB

We will use our shotgun data again since the reference genomes are also included
in the SLIMM database. For real data you will want to align against the full
SLIMM database of ~15K genomes.

*Download instructions will come soon :/*

For now we will only use the SLIMM taxonomy mapping database which is delivered with
mbtools as well.

```{r}
slimm_db <- system.file("extdata/ABVF_SP_CMP_genomes.sldb",
                        package = "mbtools")
```

which will save the db in the folder `refs`.

## Aligning shotgun reads

One advantage of SLIMM is that it is relatively agnostic to the alignment program. So
we can use our normal alignment workflow.

```{r}
fi <- system.file("extdata/shotgun", package = "mbtools") %>%
      find_read_files()
ref <- system.file("extdata/genomes/zymo_mock.fna.gz",
                   package = "mbtools")

alns <- align_short_reads(
    fi,
    threads = 3,
    reference = ref,
    use_existing = FALSE)
```

## Using SLIMM

SLIMM is again wrapped by a workflow so we can use it passing in an alignment
artifact and a configuration.

```{r}
config <- config_slimm(
    threads = 3,
    database = slimm_db
)

config
```

With this we can perform the SLIMM workflow.

```{r}
sl <- slimm(alns, config)
```

The artifact now contains the species level lineage calls...

```{r}
sl$abundance
```

Where we see no successful calls since our coverage is way to low for the example data.

However we can investigate which strains the data aligned to.

```{r}
sl$coverage[, uniqueN(strain)] %>% print()
sl$coverage
```

Where we see we actually identified the correct 10 strains. Just the coverage
is really low.

```{r}
hist(sl$coverage$reads[1][[1]])
```

Where we see that the majority of that genome has not been mapped.
