---
title: "Reference counting"
output:
    rmarkdown::html_vignette:
        toc: true
        toc_depth: 3
vignette: >
  %\VignetteIndexEntry{pkgdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Reference counting

One basic anaylisis we want to perform is to count the abundance of some references
in our data. Here reference is a broad term and may refer to transcripts, genomes,
or amplicon sequences contained in a reference database.

For this `mbtools` implements a general purpose expectation maximization (EM) counter.
It is implemented in C++ to perform fast and usually reading the alignment file takes
longer than the actual analysis.

```{r}
library(mbtools)
```

## Alignment

Let's start by aligning our short read data to a reference database of 10 microbial
genomes.

```{r}
fi <- system.file("extdata/shotgun", package = "mbtools") %>%
      find_read_files()
ref <- system.file("extdata/genomes/zymo_mock.fna.gz",
                   package = "mbtools")

alns <- align_short_reads(
    fi,
    threads = 3,
    reference = ref,
    use_existing = FALSE)
```

## EM counting

Counting is yet again a workflow and requires an alignment artifact and a configuration.

```{r}
config <- config_count(
    reference = system.file("extdata/genomes/zymo_mock.fna.gz",
                            package = "mbtools"),
    threads = 1,
    weights = TRUE
)

config
```

And we can proceed to counting:

```{r}
cn <- count_references(alns, config)
```

This creates a count artifact which contains the counts.

```{r}
cn$counts
```

In each sample we have 10 microbes with the same abundance.

```{r}
ggplot(cn$counts, aes(x=counts, y=reference)) +
    geom_point() + xlim(0, 1000) +
    facet_wrap(~ sample) + theme_minimal()
```

Looks ok, but there is a lot of noise since we have very low depth.

```{r}
ggplot(cn$counts, aes(x = counts)) + geom_histogram(bins = 8)
```

We could also compare that with a naive counting method that just uses the best
alignment score. In our case that will perform equally since we have little
multi-mapping here.

```{r}
config$method <- "naive"
cn2 <- count_references(alns, config)

counts <- cn$counts[cn2$counts, on = c("sample", "reference", "effective_length")]

ggplot(counts, aes(x = counts, y = i.counts)) +
    geom_point() + geom_smooth(method = "lm") +
    labs(x = "em", y = "naive") + theme_classic()
```

As we see they both give the same results for that simple case.
