---
title: "9 - Plate Layouts"
output:
    rmarkdown::html_vignette:
        toc: true
        toc_depth: 3
vignette: >
  %\VignetteIndexEntry{pkgdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Before even having sequencing data you will have to extract DNA or RNA from possibly many samples. This is often
done on 96-well plates which requires designing layouts for many plates. `mbtools` provides a few helpers for that.

The only thing you will need is yet again a sample manifest. This will need to have an id column again. 
Let's make one for a somewhat larger experiment.

```{r}
library(mbtools)

manifest <- data.table(
    id = paste0("S", 1:640),
    age = runif(640, 18, 97) %>% ceiling(),
    treatment = c("control", "inulin")[(runif(640) > 0.5) + 1],
    sex = c("F", "M")[(runif(640) > 0.5) + 1]
)

head(manifest)
```

You can now get plate layouts by using `layout`.

```{r}
lo <- layout(manifest)
```

This will annotate your manifest with wells and also add in blanks (those have empty IDs):

```{r}
lo$manifest
```

This will create a plate map as well. You can control the number of columns with the `ncol` argument.

```{r, fig.width = 14, fig.height = 14}
layout(manifest, ncol = 2)$layout
```

You can control the frequency of blanks with the `blank_step` argument. Set it to `Inf` or `NA` to get no blanks.

```{r, fig.width = 18, fig.height = 12}
layout(manifest, blank_step = 9, ncol = 3)$layout
```

Depending on how you like to arrange your samples you could also do so by column:

```{r, fig.width = 18, fig.height = 12}
layout(manifest, blank_step = 10, by = "col", ncol = 3)$layout
```