---
title: "Preprocessing of sequencing data"
author: "Christian Diener"
output:
    rmarkdown::html_vignette:
        toc: true
        toc_depth: 3
vignette: >
  %\VignetteIndexEntry{pkgdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Preprocessing of raw data

For filtering and trimming of the raw reads we usually use the DADA2 functions
but wrap them in a reproducible workflow step.

```{r}
library(mbtools)
```

## Finding your files

We will again use our helper function to get a list of sequencing files.

```{r}
path <- system.file("extdata/example", package = "mbtools")
files <- find_read_files(path)
print(files)
```

## Configuration

All `mbtools` workflow step come with corresponding `config_*` that returns
an example/default configuration. Changes can be done a-posteriori or by
directly passing in the parameters. We will specify a temporary directory as
storage point for the preprocessed data and truncate the forward reads to
240 bp and the reverse reads to 200 bp (based on our previous quality
assessment).

```{r}
config <- config_preprocess(out_dir = tempdir(), truncLen = c(240, 200))
config
```

We can see that there are some more parameters that we could specify.

## Running the preprocessing step

We can now run our preprocessing step.

```{r}
filtered <- preprocess(files, config)
```

This will report the percentage of passed reads on the logging interface
but you can also inspect that in detail by

```{r}
filtered$passed
```
